\vspace{-15mm}\fontsize{21pt}{10pt}\selectfont\textbf{Wordy Manuscript Title: \\Even Wordier Manuscript Subtitle\thanks{Draft manuscript. Please do not cite without the authors' permission.}}

### 2021-01-11 sections/abstract.tex
Merge with files from overleaf

\begin{abstract}
  {
    \noindent
    Microbes tend to organize into communities consisting of hundreds of species involved in complex interactions with each other.
    16S ribosomal RNA (16S rRNA) amplicon profiling provides snapshots that reveal the phylogenies and abundance profiles of these microbial communities.
    These snapshots, when collected from multiple samples, have the potential to reveal which microbes co-occur, providing a glimpse into the network of associations in these communities.
    The inference of networks from 16S data is prone to statistical artifacts.
    There are many tools for performing each step of the 16S analysis workflow, but the extent to which these steps affect the final network is still unclear.
    In this study, we perform a meticulous analysis of each step of a pipeline that can convert 16S sequencing data into a network of microbial associations.
    Through this process, we map how different choices of algorithms and parameters affect the co-occurrence network, and estimate steps which contribute most significantly to the variance. We further determine the tools and parameters that generate the most accurate and robust co-occurrence networks, based on comparison with mock and synthetic datasets.
    Ultimately, we develop a standardized pipeline that follows these default tools and parameters, but that can also help explore the outcome of any other combination of choices.
    We envisage that this pipeline for processing 16S sequencing data into networks of microbial co-occurrences could be used for integrating multiple data-sets, and for generating comparative analyses and consensus networks that can help understand and control microbial community assembly in different biomes.
    The pipeline is available as a Python package at \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE}
  }
\end{abstract}

@@

>>>>

\begin{abstract}
  {
    \noindent
    Microbes tend to organize into communities consisting of hundreds of species involved in complex interactions with each other.
    16S ribosomal RNA (16S rRNA) amplicon profiling provides snapshots that reveal the phylogenies and abundance profiles of these microbial communities.
    These snapshots, when collected from multiple samples, have the potential to reveal which microbes co-occur, providing a glimpse into the network of associations in these communities.
    The inference of networks from 16S data is prone to statistical artifacts.
    There are many tools for performing each step of the 16S analysis workflow, but the extent to which these steps affect the final network is still unclear.
    In this study, we perform a meticulous analysis of each step of a pipeline that can convert 16S sequencing data into a network of microbial associations.
    Through this process, we map how different choices of algorithms and parameters affect the co-occurrence network and estimate steps that contribute most significantly to the variance. We further determine the tools and parameters that generate the most accurate and robust co-occurrence networks based on comparison with mock and synthetic datasets.
    Ultimately, we develop a standardized pipeline (available at \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE}) that follows these default tools and parameters, but that can also help explore the outcome of any other combination of choices.
    We envisage that this pipeline could be used for integrating multiple data-sets, and for generating comparative analyses and consensus networks that can help understand and control microbial community assembly in different biomes.
}  
\end{abstract}

@@

\section*{Importance}

To understand and control the mechanisms that determine the structure and function of microbial communities, it is important to map the interrelationships between its constituent microbial species. The surge in the high-throughput sequencing of microbial communities has led to the creation of thousands of datasets containing information about microbial abundances. These abundances can be transformed into networks of co-occurrences across multiple samples, providing a glimpse into the structure of microbiomes. However, processing these datasets to obtain co-occurrence information relies on several complex steps, each of which involves multiple choices of tools and corresponding parameters. These multiple options pose questions about the accuracy and uniqueness of the inferred networks. In this study, we address this workflow and provide a systematic analysis of how these choices of tools and parameters affect the final network, and on how to select those that are most appropriate for a particular dataset.

### 2021-01-11 sections/acknowledgments.tex
Merge with files from overleaf

We are grateful to members of the Segrè lab for helpful discussions and for feedback on the manuscript. This work was partially funded by grants from the National Institutes of Health (National Institute of General Medical Sciences, award R01GM121950; National Institute of Dental and Craniofacial Research, award number R01DE024468; and National Institute on Aging, award number UH2AG064704), the U.S. Department of Energy, Office of Science, Office of Biological \& Environmental Research through the Microbial Community Analysis and Functional Evaluation in Soils SFA Program (m-CAFEs) under contract number DE-AC02-05CH11231 to Lawrence Berkeley National Laboratory, the National Science Foundation (grant NSFOCE-BSF 1635070), the Human Frontiers Science Program (RGP0020/2016), and the Boston University Interdisciplinary Biomedical Research Office.

>>>>

We are grateful to members of the Segrè lab for helpful discussions and for feedback on the manuscript. This work was partially funded by grants from the National Institutes of Health (National Institute of General Medical Sciences, award R01GM121950; National Institute of Dental and Craniofacial Research, award number R01DE024468; and National Institute on Aging, award number UH2AG064704), the U.S. Department of Energy, Office of Science, Office of Biological \& Environmental Research through the Microbial Community Analysis and Functional Evaluation in Soils SFA Program (m-CAFEs) under contract number DE-AC02-05CH11231 to Lawrence Berkeley National Laboratory, the National Science Foundation (grants 1457695 and NSFOCE-BSF 1635070), the Human Frontiers Science Program (RGP0020/2016), and the Boston University Interdisciplinary Biomedical Research Office.
KSK was supported by Simons Foundation Grant \#409704, by the Research Corporation for Science Advancement through Cottrell Scholar Award \#24010, by the Scialog grant \#26119, and by the Gordon and Betty Moore Foundation grant \#6790.08.

### 2021-01-11 sections/coverpage.tex
Merge with files from overleaf

  \vspace{1.5 cm}
  Dileep Kishore\textsuperscript{1,2},
  Gabriel Birzu\textsuperscript{3},
  Zhenjun Hu\textsuperscript{1},
  Charles DeLisi\textsuperscript{1,3},
  Kirill Korolev\textsuperscript{$\dagger$1,2,3},\\
  Daniel Segr\`{e}\textsuperscript{$\dagger$1,2,4,5}\\
  \vspace{1cm}
  \textsuperscript{1}Bioinformatics Program, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{2}Biological Design Center, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{3}Department of Physics, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{4}Department of Biomedical Engineering, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{5}Department of Biology, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{$\dagger$}Correspondence should be sent to \href{mailto:korolev@bu.edu}{korolev@bu.edu} or \href{mailto:dsegre@bu.edu}{dsegre@bu.edu}\\

>>>>

  \vspace{1.5 cm}
  Dileep Kishore\textsuperscript{1,2},
  Gabriel Birzu\textsuperscript{3,6},
  Zhenjun Hu\textsuperscript{1},
  Charles DeLisi\textsuperscript{1,3},
  Kirill S. Korolev\textsuperscript{$\dagger$1,2,3},\\
  Daniel Segr\`{e}\textsuperscript{$\dagger$1,2,4,5}\\
  \vspace{1cm}
  \textsuperscript{1}Bioinformatics Program, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{2}Biological Design Center, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{3}Department of Physics, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{4}Department of Biomedical Engineering, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{5}Department of Biology, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{6}Department of Applied Physics, Stanford University, Stanford, California, USA\\
  \textsuperscript{$\dagger$}Correspondence should be sent to \href{mailto:korolev@bu.edu}{korolev@bu.edu} or \href{mailto:dsegre@bu.edu}{dsegre@bu.edu}\\

### 2021-01-11 sections/discussion.tex
Merge with files from overleaf

The choice of taxonomy database was found to be the most important factor in the inference of a microbial co-occurrence network, contributing $\sim20\%$ of the total variance.
The frequent changes in the taxonomy nomenclature coupled with the frequency of updates to the various 16S reference databases create inherent differences \cite{Balvociute2017} in taxonomy hierarchies in these databases.
Our analysis revealed that no particular reference database performs better than the others across all scenarios. We suggest that that choice of the database should be made based on possible reported or inferred biases in the representation of given biomes in a specific databases \cite{Balvociute2017}.
The default reference database in the pipeline is the \ac{ncbi} 16S RefSeq database as it is more frequently updated and is most compatible with the blast+ query tool.
We also enable users to use custom databases \cite{Griffen2011,Ritari2015} with the blast and naive bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).

>>>>

The choice of taxonomy database was found to be the most important factor in the inference of a microbial co-occurrence network, contributing $\sim20\%$ of the total variance.
The frequent changes in the taxonomy nomenclature coupled with the frequency of updates to the various 16S reference databases create inherent differences \cite{Balvociute2017} in taxonomy hierarchies in these databases.
Our analysis revealed that no particular reference database performs better than the others across all scenarios. We suggest that that choice of the database should be made based on possible reported or inferred biases in the representation of given biomes in a specific databases \cite{Balvociute2017}.
The default reference database in the pipeline is the \ac{ncbi} 16S RefSeq database as it is more frequently updated and is most compatible with the blast+ query tool.
We also enable users to use custom databases \cite{Ritari2015} with the blast+ and naive bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).

### 2021-01-11 sections/introduction.tex
Merge with files from overleaf

 This wealth of information on the composition and functions of a community at different times and under different environmental conditions has the potential to help us understand how communities assemble and operate.
 A powerful tool for translating microbiome data into knowledge is the construction of possible inter-dependence networks across species.
 The importance of these networks of relationships is two fold: first, such networks can serve as maps that help identify hubs of keystone species \cite{Menon2018,Rottjers2018}, or basic microbiome changes that occur as a consequence of environmental perturbations or underlying host conditions \cite{Gilbert2016}; second, networks of inter-dependencies can serve as a key bridge towards building mechanistic models of microbial communities, greatly enhancing our capacity to understand and control them.
 For example, multiple studies have shown the importance of specific microbial interactions in the healthy microbiome \cite{Lloyd-Price2016} and others have shown how changes in these interactions can lead to dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
 In the context of terrestrial biogeochemistry, co-occurrence networks have proposed as a valuable approach towards reconstructing the processes leading to microbiome assembly \cite{Fierer2017}, and understanding the response of microbial communities to environmental perturbations \cite{Jiao2019}.

 Direct high-throughput measurement of interactions, e.g. through co-culture microdroplet experiments \cite{Hsu2019,Jian2020}, or spatial visualization of natural communities \cite{Wilbert2020} is possible, but it requires specific technological capabilities, and has not been used very extensively yet.
 In parallel, sequencing data across multiple samples can be used for estimating co-occurrence relationships between taxa.
 While the the relationship between directly measured interactions and statistically inferred co-occurrence is still poorly understood \cite{Zuniga2017}, a significant amount of effort has gone into estimating correlations from large microbiome sequence datasets.
 Co-occurrence networks have microbial taxa as nodes, and edges that represent the frequent co-occurrence (or negative correlations) across different datasets.

One of the most frequently used avenues for inferring co-occurrence networks is the parsing and analysis of 16S sequencing data \cite{Rottjers2018,Friedman2012}. A large number of software tools and pipelines have been developed to analyze 16S sequencing data, often focused on addressing the many known limitations of this methodology, including resolution, sequencing depth, compositional nature, sequencing errors and copy number variations \cite{Bharti2019,Pollock2018}. Popular methods for different phases of the analysis of 16S data include tools for: (i) denoising and clustering sequencing reads \cite{Caporaso2010,Callahan2016}; (ii) assigning taxonomy to the reads \cite{DeSantis2006,Quast2012}; (iii) processing and transforming the taxonomy count matrices \cite{Weiss2015}; and (iv) inferring the co-occurrence network \cite{Cougoul2019,Kurtz2015}.
Different specific algorithms are often aggregated into popular platforms (like MG-RAST \cite{Keegan2016}, Qiita \cite{qiita}) and packages (such as QIIME \cite{Caporaso2010}) that provide pipelines for 16S data analysis.
The different methods and tools developed to solve issues arising in 16S analysis can lead to vastly different inferences of community compositions and co-occurrence networks \cite{Golob2017,Weiss2016}, making it difficult to reliably compare networks across different publications and studies.
 This is partially due to the fact that existing platforms are typically focused on \ac{otu} generation and not on the effects of upstream statistical methods on the inferred co-occurrence networks.
 Furthermore, no organized framework currently exist to systematically analyze and compare existing components of the data-analysis from amplicons to networks.
  More broadly, given the lack of comprehensive comparisons between directly observed microbial interactions (e.g. from co-culture experiments) and co-occurrence networks, there is no straightforward way to determine which set of tools or methods generate the most accurate networks.

 In this study, we present a standardized 16S data-analysis pipeline called \ac{micone} that produces robust and reproducible co-occurrence networks from community 16S sequence data, and allow users to interactively explore how the network would change upon using different alternative tools and parameters at each step.
  % TODO: Link or describe MIND here (?)
  Our pipeline is coupled to an online integrative tool for the organization, visualization and analysis of inter-microbial networks.
  In addition to making this tool freely available, we implemented a systematic comparative analysis to determine which steps of the pipeline have the largest influence on the final network, and what choice seems to provide best agreement with the tested mock and synthetic datasets.
  We believe that these steps will ensure better reproducibility and easier comparison of co-occurrence networks across data-sets.
  We expect that our tool will also be useful for benchmarking future alternative methods, and for ensuring a transparent evaluation of the possible biases introduced by the use of specific tools.

>>>>

 This wealth of information on the composition and functions of a community at different times and under different environmental conditions has the potential to help us understand how communities assemble and operate.
 A powerful tool for translating microbiome data into knowledge is the construction of possible inter-dependence networks across species.
 The importance of these networks of relationships is two fold: first, such networks can serve as maps that help identify hubs of keystone species \cite{Menon2018,Rottjers2018}, or basic microbiome changes that occur as a consequence of environmental perturbations or underlying host conditions \cite{Gilbert2016}; second, networks of inter-dependencies can serve as a key bridge towards building mechanistic models of microbial communities, greatly enhancing our capacity to understand and control them.
 For example, multiple studies have shown the importance of specific microbial interactions in the healthy microbiome \cite{Lloyd-Price2016} and others have shown how changes in these interactions can lead to dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
 In the context of terrestrial bio-geochemistry, co-occurrence networks have been proposed as a valuable approach towards reconstructing the processes leading to microbiome assembly \cite{Fierer2017}, and understanding the response of microbial communities to environmental perturbations \cite{Jiao2019}.

 Direct high-throughput measurement of interactions, e.g. through co-culture micro-droplet experiments \cite{Hsu2019,Jian2020}, or spatial visualization of natural communities \cite{Wilbert2020} is possible, but it requires specific technological capabilities, and has yet to be extensively used.
 In parallel, sequencing data across multiple samples can be used for estimating co-occurrence relationships between taxa.
 While the the relationship between directly measured interactions and statistically inferred co-occurrence is still poorly understood \cite{Zuniga2017}, a significant amount of effort has gone into estimating correlations from large microbiome sequence datasets.
 Co-occurrence networks have microbial taxa as nodes, and edges that represent the frequent co-occurrence (or negative correlations) across different datasets.

One of the most frequently used avenues for inferring co-occurrence networks is the parsing and analysis of 16S sequencing data \cite{Rottjers2018,Friedman2012}. A large number of software tools and pipelines have been developed to analyze 16S sequencing data, often focused on addressing the many known limitations of this methodology, including resolution, sequencing depth, compositional nature, sequencing errors and copy number variations \cite{Bharti2019,Pollock2018}. Popular methods for different phases of the analysis of 16S data include tools for: (i) denoising and clustering sequencing reads \cite{Caporaso2010,Callahan2016}; (ii) assigning taxonomy to the reads \cite{DeSantis2006,Quast2012}; (iii) processing and transforming the taxonomy count matrices \cite{Weiss2015}; and (iv) inferring the co-occurrence network \cite{Cougoul2019,Kurtz2015}.
Different specific algorithms are often aggregated into popular platforms (like MG-RAST \cite{Keegan2016}, Qiita \cite{qiita}) and packages (such as QIIME \cite{Caporaso2010}) that provide pipelines for 16S data analysis.
The different methods and tools developed to solve issues arising in 16S analysis can lead to vastly different inferences of community compositions and co-occurrence networks \cite{Golob2017,Weiss2016}, making it difficult to reliably compare networks across different publications and studies.
 This is partially due to the fact that existing platforms are typically focused on \ac{otu} generation and not on the effects of upstream statistical methods on the inferred co-occurrence networks.
 Furthermore, no organized framework currently exist to systematically analyze and compare existing components of the data analysis from amplicons to networks.
  More broadly, given the lack of comprehensive comparisons between directly observed microbial interactions (e.g. from co-culture experiments) and co-occurrence networks, there is no straightforward way to determine which set of tools or methods generate the most accurate networks.

 In this study, we present a standardized 16S data analysis pipeline called \ac{micone} that produces robust and reproducible co-occurrence networks from community 16S sequence data, and allow users to interactively explore how the network would change upon using different alternative tools and parameters at each step.
  % TODO: Link or describe MIND here (?)
  Our pipeline is coupled to an online integrative tool for the organization, visualization and analysis of inter-microbial networks.
  In addition to making this tool freely available, we implemented a systematic comparative analysis to determine which steps of the pipeline have the largest influence on the final network, and what choice seems to provide best agreement with the tested mock and synthetic datasets.
  We believe that these steps will ensure better reproducibility and easier comparison of co-occurrence networks across datasets.
  We expect that our tool will also be useful for benchmarking future alternative methods, and for ensuring a transparent evaluation of the possible biases introduced by the use of specific tools.

### 2021-01-11 sections/methods.tex
Merge with files from overleaf

  The \ac{micone} pipeline provides both a Python API as well as a command-line interface and only requires a single configuration file.
  The configuration file lists the inputs, output and the steps to be performed during runtime, along with the parameters to be used (if different from defaults) for the various steps.
  Since the entire pipeline run-through is stored in the form of a text file (the configuration file), subsequent runs are highly reproducible and changes can be easily tracked using version control.
  It uses the nextflow workflow manager~\cite{Tommaso2015} under the hood, making it readily usable on local machines, cluster or cloud with minimal configuration change.
  It also allows for automatic parallelization of all possible processes, both within and across samples.
  The pipeline is designed to be modular: each tool or method is organized into modules which can be easily modified or replaced.
  This modular architecture simplifies the process of adding new tools (refer to modules section in the \ac{micone} documentation).
  In addition to the Python package, the entire pipeline has been containerized into a Docker~\cite{Merkel1994} image (\hl{dockerhub link}) for easy deployment and setup.
  The main components of the pipeline are detailed in the subsequent sections.

@@

  In other words, $N$ is the merged table, each column $N_i$ is the vector representation of one of the networks, and each row $L_i$ represents the one particular edge in all networks (assigned 0 if the edge does not exist in the network).

>>>>

  The \ac{micone} pipeline provides both a Python API as well as a command-line interface and only requires a single configuration file.
  The configuration file lists the inputs, output and the steps to be performed during runtime, along with the parameters to be used (if different from defaults) for the various steps.
  Since the entire pipeline run-through is stored in the form of a text file (the configuration file), subsequent runs are highly reproducible and changes can be easily tracked using version control.
  It uses the nextflow workflow manager~\cite{Tommaso2015} under the hood, making it readily usable on local machines, cluster or cloud with minimal configuration change.
  It also allows for automatic parallelization of all possible processes, both within and across samples.
  The pipeline is designed to be modular: each tool or method is organized into modules which can be easily modified or replaced.
  This modular architecture simplifies the process of adding new tools (refer to modules section in the \ac{micone} documentation).
%   In addition to the Python package, the entire pipeline has been containerized into a Docker~\cite{Merkel1994} image (\hl{dockerhub link}) for easy deployment and setup.
  The main components of the pipeline are detailed in the subsequent sections.

@@

  In other words, $N$ is the merged table, each column $N_i$ is the vector representation of one of the networks, and each row $L_i$ represents one particular edge in all networks (assigned 0 if the edge does not exist in the network).

### 2021-01-11 sections/results.tex
Merge with files from overleaf

  \begin{figure}[h]
    \centering
    \includegraphics[width=0.69\linewidth]{figure1.pdf}
    \caption{
      \textbf{The workflow of the microbial co-occurrence analysis pipeline}.
      The steps can be grouped into four major groups: \textbf{(DC)} \textbf{D}enoising and \textbf{C}lustering, \textbf{(TA)} \textbf{T}axonomy \textbf{A}ssignment, \textbf{(OP)} \textbf{O}TU or ESV \textbf{P}rocessing, and \textbf{(NI)} \textbf{N}etwork \textbf{I}nference.
      Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
      The text along the arrows describes the data that is being passed from one step to another. For details on each process and data types, see Methods.
    }
    \label{fig:figure1}
  \end{figure}

  Our analysis involves two types of data: The first type consists of sets of 16S sequencing data for real communities consisting of Stool and Oral microbiomes.
  The second are datasets synthetically or artificially created for the specific goal of helping evaluate computational analysis tools (see Methods).
  In particular, in order to objectively compare, to the extent possible, how well each step in \ac{micone} best captures the underlying data, we use both mock data (labelled mock4, mock12 and mock16) from mockrobiota~\cite{Bokulich2016} as well as, synthetically generated reads from an Illumina read simulator called ART~\cite{Huang2012}.
  These mock datasets consist of fake sequencing reads generated from reads obtained from synthetic microbial isolates mixed in know proportions. They contain the expected compositions along with the reference sequences for the organisms in the mock community.
  The synthetic reads were simulated using three different taxonomy distribution profiles, namely soil and water microbiomes obtained \ac{emp}~\cite{Thompson2017} and Stool microbiome that is used in our real community analysis~\cite{Kang2017}.
  Reference sequences were generated using \ac{ncbi} and the Decard package~\cite{Golob2017} for these taxonomy profiles.
  Detailed information on the mock communities and the settings used to generate the synthetic data are provided in the Methods section.

@@

  \begin{figure}[H]
    \centering
    \includegraphics[width=0.85\linewidth]{figure2.pdf}
  \end{figure}
  \begin{figure}[!t]
    \centering
      \caption{
      \textbf{The choice of database contributes to the most variance in the networks}.
      \textbf{(A)} The total relative variance in the networks contributed by the DC, TA and OP steps of the pipeline (right) and the linear model used to calculate the relative variance (left), see the Methods section for details.
      \textbf{(B)} All combinations of inferred networks are shown as points on a PCA plot.
      The color of the points corresponds to the taxonomy database, the shape corresponds to the denoising/clustering method and the size corresponds to whether low abundance OTUs were removed or not.
      \textbf{(B inset)} The network generated using DC=dada2, TA=GG, OP=no and NI=SPARCC and represents the particular point shown (big red square).
      The plot clearly shows that the points can be separated based on the TA step and that the differences due to the DC and OP steps are not as significant.
    }
    \label{fig:figure2}
  \end{figure}

@@

  \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figure3.pdf}
    \caption{
      \textbf{The representative sequences generated by the different denoising/clustering methods are very similar but differ in the sequences that are in low abundance.}
      \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods,
      \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences of low abundance,
      \textbf{(C, D)} The distributions of the average weighted UniFrac distance between the expected sequence profile and the calculated sequence profile in mock datasets.
      \textbf{(D)} The distributions of the average unweighted UniFrac distance show that dada2 and Deblur were the best performing methods in most of the datasets.
    }
    \label{fig:figure3}
  \end{figure}

@@

  \begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth,height=1.2\textwidth]{figure4.pdf}
  \end{figure}
  \begin{figure}[!t]
    \centering
    \caption{
      \textbf{Taxonomic reference databases vary widely in terms of their taxonomy assignments.}
      \textbf{(A)} The assignment of the top 50 representative sequences to their respective taxonomies using the three different reference databases shows how the same sequences are assigned to different Genus.
      \textbf{(B)} The percentage of \ac{otu}s assigned to the same taxonomic label when using different reference databases.
      The percentage of mismatches decrease at higher taxonomic levels but even at the Phylum level there exists around 10\% of mismatches.
      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset.
    }
    \label{fig:figure4}
  \end{figure}

@@

  For the analysis presented in this section, we used the taxonomy composition table obtained using the \ac{ncbi} reference database as the input for algorithms that infer co-occurrence associations between the microbes.
  Figure~\ref{fig:figure5}A shows the networks inferred from this dataset using the different inference algorithms.
  The different networks differ vastly in their edge-density and connectivity; even some of the edges in common to these networks have their signs inverted. Note, however, that some of these comparisons depend on the threshold that has to be applied to the pairwise correlations methods (currently 0.3, based on~\cite{Friedman2012}).
  To get a more quantitative idea, we can take a look at the nodes and edges (Figure~\ref{fig:figure5}B) in common between the networks using UpSet plots~\cite{Lex} (only \ac{magma}, \ac{mldm}, \ac{spieceasi}, \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges since they are not intended for compositional datasets).
  The results for the node intersections show that the networks have a large number of nodes in common ($63$ out of $67$ nodes in the smallest network - \ac{magma}) and no network possesses any unique node.
  The edge intersections in contrast show that only $19$ edges (out of $98$ edges in the smallest network - \ac{magma}) are in common between all the methods and each network has a large number of unique edges.
  These results indicate that there is a substantial rewiring of connections in the inferred networks.

@@

  \begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{figure5.pdf}
  \end{figure}
  \begin{figure}[t!]
    \centering
    \caption{
      \textbf{Networks generated using different network inference methods show notable differences both in terms of edge-density and connectivity}.
      \textbf{(A)} The six different networks generated by the different network inference methods are very dissimilar.
      The green links are positive associations and the orange links are negative associations.
      A threshold of 0.3 was set for the methods that infer pairwise correlations (\ac{sparcc}, Spearman, Pearson) and no threshold was set for the other methods.
      \textbf{(B)} The node overlap Upset plot~\cite{Lex} indicates that all the networks have a large number of common nodes involved in connections.
      Whereas, \textbf{(C)} The edge overlap Upset plot shows that a very small fraction of these connections are actually shared.
    }
    \label{fig:figure5}
  \end{figure}

@@

  \begin{table}[h]
    \centering
    \small
    \begin{tabular}{|c|c|c|}
      \hline
      \textbf{Process} & \textbf{Tool} & \textbf{Parameters} \\
      \hline
      Denoising and Clustering & Dada2/Deblur & default \\
      Taxonomy Assignment & \ac{ncbi} with Blast & RefSeq database \\
      OTU Processing & Based on statistical power & Dynamic cutoff \\
      Network Inference & Consensus method & - \\
      \hline
    \end{tabular}
    \caption{Default tools and parameters for the pipeline}
    \label{tab:default_options}
  \end{table}

@@

  \begin{figure}[h]
    \centering
    \includegraphics[width=0.85\linewidth]{figure6.pdf}
    \caption{
      \textbf{Network inference and taxonomic assignment have the highest influence on the inferred network structures.}
      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, TA=\ac{ncbi}, OP=on, NI=\ac{sparcc}) is compared with networks generated when one of the steps use a different tool.
      The common connections (common with the default network) are in black, connections unique to the network are colored purple and connections in the default network but not present in the current network are gray.
      \textbf{(B)} The L1 distance between the networks generated by changing one step of the default pipeline and the network generated using the default parameters.
    }
    \label{fig:figure6}
  \end{figure}

  \begin{figure}[h]
    \centering
    \includegraphics[width=17cm]{figure7.pdf}
    \caption{
      The consensus networks generated using the default pipeline settings.
      \textbf{(A)} Co-occurrence network of the Hard Palate microbiome generated from samples of healthy subjects in a periodontal diseases study.
      \textbf{(B)} Co-occurrence network of the Stool microbiome generated from samples of healthy subjects in a fecal microbiome transplant study.
  }
    \label{fig:figure7}
  \end{figure}

>>>>

  Our analysis involves two types of data: The first type consists of sets of 16S sequencing data from real communities sampled from human Stool and Oral microbiomes.
  The second are datasets synthetically or artificially created for the specific goal of helping evaluate computational analysis tools (see Methods).
  In particular, in order to objectively compare, to the extent possible, how well each step in \ac{micone} best captures the underlying data, we use both mock data (labelled mock4, mock12 and mock16) from mockrobiota~\cite{Bokulich2016} as well as, synthetically generated reads from an Illumina read simulator called ART~\cite{Huang2012}.
  These mock datasets consist of fake sequencing reads generated from reads obtained from synthetic microbial isolates mixed in know proportions. They contain the expected compositions along with the reference sequences for the organisms in the mock community.
  The synthetic reads were simulated using three different taxonomy distribution profiles, namely soil and water microbiomes obtained \ac{emp}~\cite{Thompson2017} and Stool microbiome that is used in our real community analysis~\cite{Kang2017}.
  Reference sequences were generated using \ac{ncbi} and the Decard package~\cite{Golob2017} for these taxonomy profiles.
  Detailed information on the mock communities and the settings used to generate the synthetic data are provided in the Methods section.

@@

@@

@@

@@

  For the analysis presented in this section, we used the taxonomy composition table obtained using the \ac{ncbi} reference database as the input for algorithms that infer co-occurrence associations between the microbes.
  Figure~\ref{fig:figure5}A shows the networks inferred from this dataset using the different inference algorithms.
  The different networks differ vastly in their edge-density and connectivity; even some of the edges in common to these networks have their signs inverted. Note, however, that some of these comparisons depend on the threshold that has to be applied to the pairwise correlations methods (currently 0.3, based on~\cite{Friedman2012}).
  To get a more quantitative picture of the differences between the inferred networks, we checked the distribution of common nodes and edges (Figure~\ref{fig:figure5}B) using UpSet plots~\cite{Lex} (only \ac{magma}, \ac{mldm}, \ac{spieceasi}, \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges since they are not intended for compositional datasets).
  The results for the node intersections show that the networks have a large number of nodes in common ($63$ out of $67$ nodes in the smallest network - \ac{magma}) and no network possesses any unique node.
  The edge intersections in contrast show that only $19$ edges (out of $98$ edges in the smallest network - \ac{magma}) are in common between all the methods and each network has a large number of unique edges.
  These results indicate that there is a substantial rewiring of connections in the inferred networks.

@@

@@

@@

### 2021-01-11 sections/supplementary.tex
Add supplementary table
Update supplementary table
Merge with files from overleaf

  \hl{Copy paste table here once done}

@@

  \begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{figureS1.pdf}
  \caption{
    \textbf{Comparison of various denoising and clustering algorithms used in the pipeline}.
    (A, B) Correlation ofjk the abundances of the taxa that are in common between the count matrices created by two different methods.
    (A) The best correlation (most similar methods) is between open-reference and denovo.
    (B) The worst correlation (least similar methods) is between open-reference and dada2.
    (C) A heatmap showing the $\mathrm{R}^2$ of all pairwise comparisons of the methods.
  }
  \label{fig:figureS1}
\end{figure}

@@

  \begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figureS6.pdf}
    \caption{
      Calculation of presence threshold that is applied on the OTU table in the OTU processing (OP) step of the pipeline.
      This presence threhold $p_t$ is dependent on the number of samples in the dataset and the required correlation stength threshold.
  }
    \label{fig:figureS6}
  \end{figure}

@@

  \begin{figure}[h]
    \centering
    \includegraphics[width=0.8\linewidth]{pdf/denoise_network.pdf}
    \caption{A network showing union and intersection of networks generated using certain combination of methods}
    \label{fig:figureS5}
  \end{figure}

>>>>

@@

  \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{table}{0}

  \input{sections/supplementary_table.tex}

  \begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{figureS1.pdf}
  \caption{
    \textbf{Comparison of various denoising and clustering algorithms used in the pipeline}.
    (A, B) Correlation of the abundances of the taxa that are in common between the count matrices created by two different methods.
    (A) The worst correlation (least similar methods) is between open-reference and dada2.
    (B) The best correlation (most similar methods) is between open-reference and denovo.
    (C) A heatmap showing the $\mathrm{R}^2$ of all pairwise comparisons of the methods.
  }
  \label{fig:figureS1}
\end{figure}

@@

%   \begin{figure}[h]
%     \centering
%     \includegraphics[width=\linewidth]{figureS6.pdf}
%     \caption{
%       Calculation of presence threshold that is applied on the OTU table in the OTU processing (OP) step of the pipeline.
%       This presence threhold $p_t$ is dependent on the number of samples in the dataset and the required correlation stength threshold.
%   }
%     \label{fig:figureS6}
%   \end{figure}

@@

  \begin{figure}[h]
    \centering
    \includegraphics[width=0.8\linewidth]{pdf/denoise_network.pdf}
    \caption{A network showing union (A) and intersection (B) of networks generated using different denoising and clustering tools on the Stool dataset.}
    \label{fig:figureS5}
  \end{figure}
